Identifying Heterogeneity in SAR Data with New Test Statistics

### 2024-11-22 Text/2024PresentationsChina.tex
Fresh run of RMarkdown R1 file
Plots for the slides

\title{Identifying Heterogeneity in SAR Data with New Test Statistics}
\subtitle{There are no too old problems}
\author{Alejandro C.\ Frery}
\institute{Victoria University of Wellington\\
	School of Mathematics and Statistics\\
	New Zealand}
\date{November, 2024}

@@

	\begin{frame}{Starting point}
		\begin{itemize}[<+->]
			\item Synthetic Aperture Radar (SAR) technology is essential for environmental monitoring and disaster management. 
			\item It provides
			images under various conditions, including day or night and adverse weather
			situations~\citep{Moreira2013,Mu2019}. 
			\item The effective use of SAR
			data relies on understanding of their statistical properties
			because it is corrupted by speckle: a noise-like interference effect~\citep{Argenti2013}.
		\end{itemize}
	\end{frame}

	\begin{frame}[standout]
		\frametitle{Takeaways}
		You may forget about the technical contents of this talk, but remember:
		\begin{itemize}[<+->]
			\item There are no ``too old'' problems that cannot be revisited.
			\item The data statistical properties always bring valuable information.
			\item A careful mathematical manipulation often leads to new exciting results.
		\end{itemize}
	\end{frame}

>>>>

\title{Identifying Heterogeneity in SAR Data with New Test Statistics}
\subtitle{No problem is too old, and you don't always need deep learning}
\author{Alejandro C.\ Frery}
\institute{Victoria University of Wellington\\
School of Mathematics and Statistics\\
New Zealand}
\date{November, 2024}

@@

\begin{frame}{Starting point}
\begin{itemize}[<+->]
	\item Synthetic Aperture Radar (SAR) technology is essential for environmental monitoring and disaster management. 
	\item It provides
	images under various conditions, including day or night and adverse weather
	situations~\citep{Moreira2013,Mu2019}. 
	\item The effective use of SAR
	data relies on understanding of their statistical properties
	because it is corrupted by speckle: a noise-like interference effect~\citep{Argenti2013}.
	\item Speckle in intensity format is non-Gaussian. 
	\item The
	\(\mathcal{G}^0\) distribution is suitable for SAR data and includes
	the Gamma law as the limiting case for fully-developed
	speckle~\citep{Ferreira2020}.
	\item We improve the identification of potential roughness
	features in SAR intensity data, i.e., departures from the Gamma distribution (fully-developed
	speckle).
\end{itemize}
\end{frame}

\begin{frame}{The problem}
\centering
\includegraphics[width=55mm]{../Figures/PNG/Rotterdam_1024}\hskip1em
\includegraphics[width=55mm]{../Figures/PNG/lake_512}

Apart from the brightness, notice that different types of areas have distinct roughness.
\end{frame}

\begin{frame}{Statistical models}
The main models are the Gamma (suitable for fully-developed speckle) and
\(\mathcal{G}_I^0\) (able to describe roughness) distributions~\citep{Frery1997}.

They are characterized by the probability
density functions (pdfs): 
\begin{align}
	f_Z(z;L, \mu\mid \Gamma_{\text{SAR}})&=\frac{L^L}{\Gamma(L)\mu^L}z^{L-1}\exp\left\{-Lz/\mu\right\} \mathbbm 1_{\mathbbm R_+}(z)\label{E:gamma1}\\
	\intertext{ and }
	f_Z(z; \alpha, \gamma, L \mid \mathcal{G}_I^0)&=\frac{L^L\Gamma(L-\alpha)}{\gamma^{\alpha}\Gamma(-\alpha)\Gamma(L)}\cdot\frac{z^{L-1}}{(\gamma+Lz)^{L-\alpha}} \mathbbm 1_{\mathbbm R_+}(z),\label{E:gi01}
\end{align} 
where \(\mu > 0\) is the mean, \(\gamma > 0\) is the scale,
\(\alpha < 0\) measures the roughness, \(L \geq 1\) is the number of
looks, \(\Gamma(\cdot)\) is the gamma function, and
\(\mathbbm 1_{A}(z)\) is the indicator function of the set \(A\).
\end{frame}

\begin{frame}{The statistical approach}
\begin{columns}
	\begin{column}{.63\linewidth}
{\includegraphics<1>[width=\linewidth]{../Figures/PDF/GammaGI0Densities}}
{\includegraphics<2>[width=\linewidth]{../Figures/PDF/GammaGI0DensitiesSemilog}}
	\end{column}
	\begin{column}{.3\linewidth}
	We need to identify the model that best describes the data.

	But it is a challenging task because
	\begin{itemize}
		\item Samples are small, and
		\item Parameter estimators are tricky.
	\end{itemize}
	\end{column}
\end{columns}
\end{frame}

\begin{frame}[standout]
\frametitle{Takeaways}
You may forget about the technical contents of this talk, but remember:
\begin{itemize}[<+->]
	\item No problem is too old to be revisited.
	\item The data statistical properties always bring valuable information.
	\item A careful mathematical manipulation often leads to new exciting results.
	\item Reproducibility goes a long way!
\end{itemize}
\end{frame}
